---
title: "Compare fitted parameters from weighted sites vs. not"
output: html_document
date: "2024-11-14"
---

# Experiment Overview

Parameters were estimated by the decemedip model on 3 healthy plasma samples (each has low, medium and high read coverage).

This markdown shows difference made by applying uniform and non-uniform likelihood weights on reference sites

- Uniformly weighted (i.e., unweighted):
  - Weight on cell type-specific sites: 1
  - Weight on anchor sites: 1

- Non-uniformly weighted (i.e., weighted):
  - Weight on cell type-specific sites: 1
  - Weight on anchor sites: 0.5

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = '/scratch/st-kdkortha-1/nshen7/cfMeDIP_deconv/cfMeDIP-deconv-experiments')
knitr::opts_chunk$set(echo = TRUE)
setwd('/scratch/st-kdkortha-1/nshen7/cfMeDIP_deconv/cfMeDIP-deconv-experiments')
source('code/SETPATHS.R')
library(decemedip)
devtools::load_all('../decemedip/') # to load new changes in functions
```

```{r, include=FALSE}
read_dir_0 <- here('data', 'interim', 'synthetic_benchmark_new', 'benchmark_simulated_allcelltypes', '02_extract_parameter_posterior')
read_dir_1 <- here('data', 'interim', 'synthetic_benchmark_new', 'benchmark_simulated_allcelltypes_weighted', '02_extract_parameter_posterior')

ref_cts <- readRDS(here('data', 'metadata', 'references', 'MethAtlas', 
                        'summarizedexperiment_cell_type_specific_atlas_v3.rds'))
```

```{r, include=FALSE}
gatherData <- function(COVERAGE) {
  name_seg <- paste0('_cov', COVERAGE)

  w_mu_0 <- fread(here(read_dir_0, paste0('param_posterior_w_mu', name_seg, '.csv'))) |> mutate(prior_type = 'Unweighted')
  w_sigma_0 <- fread(here(read_dir_0, paste0('param_posterior_w_sigma', name_seg, '.csv'))) |> mutate(prior_type = 'Unweighted')
  pi_0 <- fread(here(read_dir_0, paste0('param_posterior_pi', name_seg, '.csv'))) |> mutate(prior_type = 'Unweighted')
  
  w_mu_1 <- fread(here(read_dir_1, paste0('param_posterior_w_mu', name_seg, '.csv'))) |> mutate(prior_type = 'Weighted')
  w_sigma_1 <- fread(here(read_dir_1, paste0('param_posterior_w_sigma', name_seg, '.csv'))) |> mutate(prior_type = 'Weighted')
  pi_1 <- fread(here(read_dir_1, paste0('param_posterior_pi', name_seg, '.csv'))) |> mutate(prior_type = 'Weighted')
  
  w_mu <- rbind(w_mu_0, w_mu_1)
  w_sigma <- rbind(w_sigma_0, w_sigma_1)
  pi <- rbind(pi_0, pi_1)
  
  return(list(w_mu = w_mu, w_sigma = w_sigma, pi = pi))
}

plotMu <- function(w_mu) {
  p <- w_mu |> 
    ggplot(aes(param_name, mean, color = prior_type)) +
    geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), 
                   position = position_dodge2(width = 0.35)) +
    geom_point(position = position_dodge2(width = 0.35)) +
    theme_classic()
  return(p)
}

plotSigma <- function(w_sigma) {
  p <- w_sigma |> 
    ggplot(aes(param_name, mean, color = prior_type)) +
    geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), 
                   position = position_dodge2(width = 0.35)) +
    geom_point(position = position_dodge2(width = 0.35)) +
    theme_classic()
  return(p)
}

plotPi <- function(pi) {
  p <- pi |> 
    mutate(cell_type = factor(cell_type, levels = colnames(ref_cts))) |>
    ggplot(aes(cell_type, mean, color = prior_type)) +
    geom_linerange(aes(ymin = `2.5%`, ymax = `97.5%`), 
                   position = position_dodge2(width = 0.35)) +
    geom_point(position = position_dodge2(width = 0.35)) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  return(p)
}

```

# Low-coverage samples

```{r}
cur_list <- gatherData(COVERAGE = 1) 
list2env(cur_list, envir = .GlobalEnv)
```

```{r}
plotMu(w_mu)
```

```{r}
plotSigma(w_sigma)
```

```{r}
plotPi(pi)
```

# Medium-coverage samples

```{r}
cur_list <- gatherData(COVERAGE = 2) 
list2env(cur_list, envir = .GlobalEnv)
```

```{r}
plotMu(w_mu)
```

```{r}
plotSigma(w_sigma)
```

```{r}
plotPi(pi)
```

# High-coverage samples

```{r}
cur_list <- gatherData(COVERAGE = 3) 
list2env(cur_list, envir = .GlobalEnv)
```

```{r}
plotMu(w_mu)
```

```{r}
plotSigma(w_sigma)
```

```{r}
plotPi(pi)
```

